some title

### 2018-01-11 2018/paper.tex
partial corr theorem

Then we need to derive a new vector $\tilde x$ with constants added to each component.
Geometrically adding a vector of costants means adding a vector of all ones
scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
By the properties of projection it is of no difference whether to project
the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
and $\beta x$ — the result is the same.
So, while $\beta x$ is projected onto a span of $x^c$, the projection of $\alpha \mathbf{1}$
onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
in Figure~\ref{fig:corr_tildex}.
Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.

@@

\begin{proof}
Firstly, we consider the case when $\hat \beta_2 > 0$ so the main picuture is of the form depicted in Figure~\ref{fig:corr_pos_base}.
It has been shown earlier that the correlation coefficient represents the angle betweem two random vectors.
So in order to complete the proof we need to find the appropriate angles and compare them.

@@

>>>>

Then we need to derive a new vector $\tilde x$ with constants added to each component.
Geometrically adding a vector of costants means adding a vector of all ones
scaled by $\alpha \in \mathbb{R}$, i.e., $\alpha \mathbf{1}$.
Then the new vector $\tilde x$ can be broken up into a sum of $\alpha \mathbf{1}$ and
$\beta x$, $\alpha, \beta \in \mathbb{R}$, which can be seen in Figure~\ref{fig:corr_tildex}.
After that we will project this new vector $\tilde x$ onto $Lin^{\perp}(\mathbf{1})$.
By the properties of projection it is of no difference whether to project
the whole vector $\tilde x$ or project its parts $\alpha \mathbf{1}$
and $\beta x$ — the result is the same.
So, while $\beta x$ is projected onto the span of $x^c$, the projection of $\alpha \mathbf{1}$
onto the orthohgonal space $Lin^{\perp}(\mathbf{1})$ yields zero as demonstrated
in Figure~\ref{fig:corr_tildex}.
Moreover, it follows that the angle between $\tilde x$ and $y$ is still $\varphi$.

@@

\begin{proof}
Firstly, we consider the case when $\hat \beta_2 > 0$ and introduce the base
picuture of the proof as depicted in Figure~\ref{fig:corr_pos_base}.
It has been shown earlier that the correlation coefficient represents the angle
betweem two random vectors.
So in order to complete the proof we need to find the appropriate angles and compare them.

@@

\subsection{Partial correlation as correlaiton between residuals}

\marginnote{
???
}

\begin{theorem}
Partial correlaiton between $X$ and $Y$ holding $Z$ fixed is the (-?)
correlaiton coefficient betwenn the residuals $\hat u$ in the regeression model
\[
X = \alpha_1 Y + \alpha_2 Z + u
\]
and the residuals $\hat v$ in the model
\[
Y = \beta_1 Z + \beta_2 Z + v
\]
\end{theorem}

\begin{proof}
The first step is to find the residuals in the mentioned regresisions.
For example, in order to get $\hat u$ we regress $X$ onto $Lin(Y,Z)$
which results in $\hat X$. Then we take the difference $X - \hat X = \hat u$
and project it as well as $X$ itself onto $Lin^{\perp}(Z)$ as demonstrated
in Figure~\ref{fig:pcorr_t_x}.

Figure~\ref{fig:pcorr_t_y} shows the same step for obtainig $\hat v$.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_x.pdf}
\label{fig:pcorr_t_x}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{images/pcorr_theorem_y.pdf}
\label{fig:pcorr_t_y}}
\caption{\subref{fig:pcorr_t_x}: $\hat u$ form regression of x onto $Y$ and $Z$, $\hat u$ projected;
\subref{fig:pcorr_t_y}: $\hat v$ from regression of $Y$ onto $X$ and $Z$, $\hat v$ projected.}
\end{center}
\end{figure}

After all we put these figure together and the goal is to find the angle between
the red and blue lines which is the same as the angle between the dashed
red and blue lines. However, before that we need to apply translating to them to
get this angle as shown in Figure~\ref{fig:pcorr_t_lin}. Finally, using the properties of
complementary angles, angles in triangles and parallelograms
we conclude that the angle between the residuals $\tilde u$ and $\tilde v$
is eqaul to $\varphi$.

\begin{figure}[ht!]
\begin{center}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{images/pcorr_t_xy.pdf}
\label{fig:pcorr_t_xy}}
%\hspace{4ex}
\subfigure[]{
\includegraphics[width=0.45\linewidth]{images/pcorr_t_lin.pdf}
\label{fig:pcorr_t_lin}}
\caption{\subref{fig:pcorr_t_x}: The residuals of both regressions;
\subref{fig:pcorr_t_y}: $Lin^{\perp}(Z)$.}
\end{center}
\end{figure}
